Combinatorics of the geometry of Wilson loop diagrams I: equivalence classes via matroids and polytopes

### 2019-07-09 GWLDCI.tex
more proofreading paper 1

Propagators are undirected, so $p = (i,j) = (j,i)$. In previous work, we have used the convention, we write $p = (i,j)$ with  $i +1 \leq j$ relative to the first vertex. \sanote{Susama stopped here}

@@

@@

\begin{dfn} \label{VPropdfn}
Let $W = (\cP, [n])$.
\begin{enumerate}
\item For $p = (i_p,j_p) \in \cP$, define $V(p) = \{i_p, i_p+1, j_p, j_p+1\}$ to be the set of vertices supporting $p$. Similarly, for $P \subseteq \cP$ we define the set $V(P) = \cup_{p \in P} V(p)$ to be the vertex support of $P$.
\item For $U \subseteq [n]$, write $\Prop(U) = \{ p \in \cP  \ | \ V(p) \cap U \neq \emptyset \} $ to denote the set of propagators which are at least partially supported on $U$.
\item For $P \subseteq \cP$, define the {\em propagator flat}\footnote{This choice of terminology is justified in Lemma~\ref{lem facts about WLD matroids} below.} $F(P) = V(P^c)^c$ to be the set of vertices in $[n]$ that do not support any propagators outside of the set $P$. Here, as throughout, we are using the superscript $c$ for the set theoretic complement.
\item The set of vertices that do not support any propagators is denoted $F(\emptyset)$. Vertices in this set are called {\em non-supporting}.  
\end{enumerate}
\end{dfn}

\begin{rmk}\label{alt F(P) rmk}
An equivalent definition of $F(P)$ is \bas F(P)  = \big(V(P) \setminus V(P^c) \big) \cup \{\text{all non-supporting vertices in $W$}\} \;,\eas 
i.e. $F(P)$ can be thought of as the set of vertices that either only support propagators in $P$ or do not support any propagators at all. Furthermore, note that by construction we have $\Prop(F(P)) \subseteq P$.  This containment is strict if there is any propagator $p$ in $P$ with the property that every vertex in the support of $p$ also supports a propagator not in $P$.
\end{rmk}

It is sometimes useful to discuss propagators in terms of the edges supporting them (rather than the vertices) so we make the following definition.

In this manner, the propagator $p = (i, j)$ is supported by the $i$th and $j$th edges.

We now define the notion of an admissible diagram, which imposes certain density and non-crossing conditions on the configuration of the propagators. The motivation for these conditions comes from the link to SYM $N=4$ theory, where the admissible Wilson loop diagrams are exactly those that correspond to $N^kMHV$ diagrams; see for example \hlfix{[ref].}{I don't think we should go into any detail here, but I wanted to emphasise that there is a reason for choosing these particular conditions.}

@@

In other words, a Wilson loop diagram, $(\cP, [n])$ is admissible only if $n \geq |\cP| +4$, and has no crossing propagators, no propagators of the form $p = (i, i+1)$, and no pairs of propagators that start and end on the same pair of non-adjacent edges.\note{Sian: can we leave this sentence out? By dropping ``iff'' to ``only if'' we make it sound like the converse isn't true.}

@@

@@

\vspace{3pt}
\begin{eg}
The Wilson loop diagram $W$ in Example~\ref{eg of WLD} is an example of an admissible diagram. It has a single non-supporting vertex, namely the vertex labeled 7, so we have $F(\emptyset) = \{7\}$. For $p = (5, 8)$, we have $V(p) = \{5,6,8,1\}$ and $F(p) = \{6, 7,8\}$: the vertices 1 and 5 are excluded from $F(p)$ since they support other propagators as well, and 7 is included (despite not being in the support of $p$) since it supports no propagators at all. Finally, the subdiagram $(\{(2, 4), (1, 4)\}, \{1, 2, 3, 4, 5, \})$ is an example of an exact subdiagram; in fact it is the only non-trivial exact subdiagram in $W$.
\end{eg}

@@

Let $W = (\cP,[n])$ and $W' = (\cP',[n])$, and say that $W$ and $W'$ are related if
\begin{enumerate}
\item There exist exact subdiagrams $(P, V(P)) \subseteq W$ and $(P', V(P')) \subseteq W'$ with $V(P) =  V(P')$.
\item The complementary subdiagrams are identical, i.e. $(\cP \setminus P, V(P)^c) = (\cP' \setminus P', V(P')^c)$.
\end{enumerate}
Let $\sim$ denote the transitive closure of this relation. This is an equivalence relation on the set of all admissible Wilson loop diagrams with vertex set $[n]$.
\end{dfn}

\begin{eg} \label{eg:equivdiags}
Note that we genuinely do need to take the transitive closure of this relation in order to obtain an equivalence relation. For example, consider the following three Wilson loop diagrams:
\bas
W_1 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
    \drawprop{5}{0}{8}{0}
    \drawprop{5}{1}{7}{0}
		\end{scope}
	\end{tikzpicture} \quad
W_2 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{2}{0}{4}{-1}
    \drawprop{5}{0}{8}{0}
    \drawprop{6}{0}{8}{-1}
		\end{scope}
	\end{tikzpicture} \quad
W_3 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawprop{1}{0}{4}{0}
	\drawprop{1}{1}{3}{0}
    \drawprop{5}{0}{8}{0}
    \drawprop{6}{0}{8}{-1}
		\end{scope}
	\end{tikzpicture} .
\eas

We have $W_1 \sim W_2$ because they are identical except on the exact subdiagrams 
\[(\{(5,8), (5,7)\}, \{5,6,7,8,1\}) \text{ and }(\{(5,8), (6,8)\}, \{5,6,7,8,1\})\] respectively. We also have $W_2 \sim W_3$ due to the exact subdiagrams 
\[(\{(1,4), (2,4)\}, \{1,2,3,4,5\}) \text{ and }(\{(1,4), (1,3)\}, \{1,2,3,4,5\}).\] 
However, the difference between $W_1$ and $W_3$ is the union of two exact subdiagrams, which is not itself exact. \end{eg}

As described in the introduction, Wilson loop diagrams come from the study of SYM $N = 4$ in the physics literature and therefore each diagram $W$ is associated to a matrix of data $C(W)$, which we now describe.

>>>>

Wherever possible, we drop the subscrips on propagators and their identifying vertices. Propagators are undirected, so $p = (i,j) = (j,i)$. In previous work, we have used the convention, we write $p = (i,j)$ with  $i +1 \leq j$ relative to the first vertex.

@@

To simplify language, we let the edges inherit a cyclic ordering from the vertices:

and thus speak of a propagator $p = (i, j)$ as being supported by the $i$th and $j$th edges. \sanote{moved/altered dfn and sandwiching paragrahs to improve flow}

@@

\begin{dfn} \label{VPropdfn}
Let $W = (\cP, [n])$.
\begin{enumerate}
\item For $p = (i,j) \in \cP$, define $V(p) = \{i, i+1, j, j+1\}$ to be the set of vertices supporting $p$. Similarly, for $P \subseteq \cP$ we define the set $V(P) = \cup_{p \in P} V(p)$ to be the vertex support of $P$.
\item For $U \subseteq [n]$, write $\Prop(U) = \{ p \in \cP  \ | \ V(p) \cap U \neq \emptyset \} $ to denote the set of propagators which are at least partially supported on $U$.
\item For $P \subseteq \cP$, define the {\em propagator flat}\footnote{This choice of terminology is justified in Lemma~\ref{lem facts about WLD matroids} below.} $F(P) = V(P^c)^c$ to be the set of vertices in $[n]$ that do not support any propagators outside of the set $P$. Here, as throughout, we are using the superscript $c$ for the set theoretic complement.
\item The set of vertices that do not support any propagators is denoted $F(\emptyset)$. Vertices in this set are called {\em non-supporting}.  
\end{enumerate}
\end{dfn}

\begin{rmk}\label{alt F(P) rmk}
An equivalent definition of $F(P)$ is \bas F(P)  = \big(V(P) \setminus V(P^c) \big) \cup \{\text{all non-supporting vertices in $W$}\} \;,\eas 
i.e. $F(P)$ can be thought of as the set of vertices that either only support propagators in $P$ or do not support any propagators at all. Furthermore, note that by construction we have $\Prop(F(P)) \subseteq P$.  This containment is strict if there is any propagator $p$ in $P$ with the property that every vertex in the support of $p$ also supports a propagator not in $P$. \sanote{whoever added this last sentence, thank you. It is an oft overlooked important point.}
\end{rmk}

We now define the notion of an admissible diagram, which imposes certain density and non-crossing conditions on the configuration of the propagators. The motivation for these conditions comes from the link to SYM $N=4$ theory, where the admissible Wilson loop diagrams are exactly those that correspond to $N^kMHV$ diagrams; see for example \cite{Amplituhedronsquared}.

@@

In other words, a Wilson loop diagram, $(\cP, [n])$ is admissible only if $n \geq |\cP| +4$, and has no crossing propagators, no propagators of the form $p = (i, i+1)$, and no pairs of propagators that start and end on the same pair of non-adjacent edges.\note{Sian: can we leave this sentence out? By dropping ``iff'' to ``only if'' we make it sound like the converse isn't true.} \sanote{no objections. Want to check that we don't use this property in the paper though. If not, lets nix}

@@

{\color{violet}Note that by Definition \ref{admisdfn}, any single propagator and its support defines an exact subdiagram. These are called trivial sub diagrams.}

@@

\vspace{3pt}
\begin{eg}
The Wilson loop diagram $W$ in Example~\ref{eg of WLD} is an example of an admissible diagram. It has a single non-supporting vertex, namely the vertex labeled 7, so we have $F(\emptyset) = \{7\}$. For $p = (5, 8)$, we have $V(p) = \{5,6,8,1\}$ and $F(p) = \{6, 7,8\}$: the vertices 1 and 5 are excluded from $F(p)$ since they support other propagators as well, and 7 is included (despite not being in the support of $p$) since it supports no propagators at all. Finally, the subdiagram $(\{(2, 4), (1, 4)\}, \{1, 2, 3, 4, 5, \})$ is an example of an exact subdiagram; \hlfix{in fact it is the only non-trivial exact subdiagram in $W$.}{Either nix or include violet text above. I lean towards nixing.}
\end{eg}

@@

Let $W = (\cP,[n])$ and $W' = (\cP',[n])$ be two admissible subdiagrams. We say that $W$ and $W'$ are related if
\begin{enumerate}
\item There exist exact subdiagrams $(P, V(P)) \subseteq W$ and $(P', V(P')) \subseteq W'$ with the same number of propagators and suppored on the same vertices; $|P| = |P'|$ and  $V(P) =  V(P')$.
\item The complementary subdiagrams (i.e. all other propagators and vertices) are identical, i.e. $(\cP \setminus P, V(P)^c) = (\cP' \setminus P', V(P')^c)$. \sanote{made some changes}
\end{enumerate}
Let $\sim$ denote the transitive closure of this relation. This is an equivalence relation on the set of all admissible Wilson loop diagrams with vertex set $[n]$.
\end{dfn}

\begin{eg} \label{eg:equivdiags}
Note that we genuinely do need to take the transitive closure of this relation in order to obtain an equivalence relation. For example, consider the following three Wilson loop diagrams: 
\bas
W_1 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawlabeledprop{1}{0}{4}{0}{\footnotesize $q$}
	\drawlabeledprop{2}{0}{4}{-1}{\footnotesize  $p$}
    \drawlabeledprop{5}{0}{8}{0}{\footnotesize \quad $r$}
    \drawlabeledprop{5}{1}{7}{0}{\footnotesize \quad $s$}
		\end{scope}
	\end{tikzpicture} \quad
W_2 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawlabeledprop{1}{0}{4}{0}{\footnotesize $q$}
	\drawlabeledprop{2}{0}{4}{-1}{\footnotesize $p$}
    \drawlabeledprop{5}{0}{8}{0}{\footnotesize \quad $u$}
    \drawlabeledprop{6}{0}{8}{-1}{\footnotesize $v$}
		\end{scope}
	\end{tikzpicture} \quad
W_3 = \begin{tikzpicture}[rotate=67.5,baseline=(current bounding box.east)]
	\begin{scope}
	\drawWLD{8}{1.5}
	\drawnumbers
	\drawlabeledprop{1}{0}{4}{0}{\footnotesize $a$}
	\drawlabeledprop{1}{1}{3}{0}{\footnotesize \quad $b$}
    \drawlabeledprop{5}{0}{8}{0}{\footnotesize \quad $u$}
    \drawlabeledprop{6}{0}{8}{-1}{\footnotesize $v$}
		\end{scope}
	\end{tikzpicture} .
\eas

We have $W_1 \sim W_2$ because they are identical except on the exact subdiagrams 
\[(\{r, s\}, \{5,6,7,8,1\}) \text{ and }(\{u, v\}, \{5,6,7,8,1\})\] respectively. We also have $W_2 \sim W_3$ due to the exact subdiagrams 
\[(\{p, q\}, \{1,2,3,4,5\}) \text{ and }(\{a, b\}, \{1,2,3,4,5\}).\] 
However, the difference between $W_1$ and $W_3$ is the union of two exact subdiagrams, which is not itself exact. \sanote{labeled propagators to make easier to read} \end{eg}

As described in the introduction, Wilson loop diagrams come from the study of SYM $N = 4$ in the physics literature \st{and therefore} \hlfix{where each propagator corresponds to a section of  a $k$-line bundle over twistor space. Thus}{I feel the strike out was too vague. The highlighted may be too technical. Can someone find baby bear's text?} each diagram $W$ is associated to a matrix of data $C(W)$, which we now describe. \sanote{Susama ended here}
